---
title: "Microfragmentation Immunity Data"
author: "Lou Schlecker & Rachel Wright"
date: "7/27/2021"
output:
  html_document: default
  pdf_document: default
---

```{r load libraries, include = FALSE}
library(tidyverse) # for data wrangling and plotting
library(MCMCglmm) # for Bayesian generalized linear mixed model statistical analysis
library(RColorBrewer)

# setwd("Z:/Lou_Fragmentation/R_test/immune_data_ready")
```

# Immunity Analyses
We have immunity data for 96 fragments

```{r Load Data, cache = TRUE, message=FALSE, warning = FALSE}
allData <- read.csv("allData.csv")
head(allData)
```

Combine immune data with the growth data

```{r combine data}
growthData <- read.csv("growthData.csv")
combinedData <- merge(allData, growthData, by=c("sam","geno"))
head(combinedData)
```

Combine Genets X and V, which we know from sequence data are clones
Rename them both "U"

```{r Combine clones growth data}
combinedData$geno <- gsub("X|V", "U", combinedData$geno)
combinedData$geno
```

Explore immune data

prot.10uL. = protein concentration in 10 µL sample included in protein assay (mg/mL) 

protN = total protein normalized to surface area (mg)
(protein concentration in 10 µL sample (mg/mL) * airbrush volume (mL)) / surface area (cm2)
final units = mg/cm2

ppo (total phenoloxidase potential) and po (active phenoloxidase) = ∆A490 / min
pox (peroxidase) = ∆A470 / min
cat (catalase) = ∆H2O2 / min

ppoN, poN, poxN, and catN = normalized to total protein
final units = change in absorbance / min per mg protein

```{r phenoloxidase histogram}

combinedData %>% select(poN, ppoN, species) %>%
  pivot_longer(cols = c(poN, ppoN),
               names_to = "measurement",
               values_to = "value") %>%
  ggplot(aes(x = value, fill = measurement)) +
  facet_grid(~ species) +
  ylab("Count") +
  xlab("Phenoloxidase Activity (∆A490 / min * protein)") +
  geom_histogram(position = "identity", alpha = 0.5, bins = 20) +
  theme_classic() +
  scale_color_manual(name = "Measurement",
                    labels = c("Active PO", "Total Potential PO"))+
  scale_fill_brewer(palette = "Set1")
```

```{r peroxidase histogram}
combinedData %>% select(poxN, species) %>%
  pivot_longer(cols = c(poxN),
               names_to = "measurement",
               values_to = "value") %>%
  ggplot(aes(x = value, fill = measurement)) +
  facet_grid(~ species) +
  ylab("Count") +
  xlab("Peroxidase Activity (∆A470 / min * protein)") +
  geom_histogram(position = "identity", alpha = 0.65, bins = 20) +
  theme_classic() +
  scale_fill_manual(values=c("steelblue"),
                      name = "Measurement",
                    labels = c("Peroxidase"))
```

```{r catalase histogram}
combinedData %>% select(catN, species) %>%
  pivot_longer(cols = c(catN),
               names_to = "measurement",
               values_to = "value") %>%
  ggplot(aes(x = value, fill = measurement)) +
  facet_grid(~ species) +
  ylab("Count") +
  xlab("Catalase Activity (∆H2O2 / min * protein)") +
  geom_histogram(position = "identity", alpha = 0.65, bins = 20) +
  theme_classic() +
  scale_fill_manual(values=c("steelblue"),
                      name = "Measurement",
                    labels = c("Catalase"))
  
```

Catalase

The values for catalase are negative because H2O2 is being consumed by catalase. Multiply by -1 to convert to "positive" values for easier interpretation.

```{r convert catalase}
combinedData <- combinedData %>% mutate(catNP = catN*-1)
```

```{r catalase histogram positive}
combinedData %>% select(catNP, species) %>%
  pivot_longer(cols = c(catNP),
               names_to = "measurement",
               values_to = "value") %>%
  ggplot(aes(x = value, fill = measurement)) +
  facet_grid(~ species) +
  ylab("Count") +
  xlab("Catalase Activity (∆H2O2 / min * protein)") +
  geom_histogram(position = "identity", alpha = 0.65, bins = 20) +
  theme_classic() +
  scale_fill_manual(values=c("steelblue"),
                      name = "Measurement",
                    labels = c("Catalase"))
```

Summarize all of the data before outliers are removed

```{r boxplot before outliers removed}
boxplot.before.outliers <- combinedData %>% select(poN, ppoN, poxN, catNP, species) %>%
  pivot_longer(cols = c(poN, ppoN, poxN, catNP),
               names_to = "measurement",
               values_to = "value") %>%
  ggplot(aes(x = species, y = value, fill = measurement)) +
  facet_wrap(~ measurement, scales = "free_y") +
  geom_boxplot() +
  theme_classic() +
  scale_fill_manual(values=c("red", "green", "orange", "purple"),
                      name = "Measurement",
                    labels = c("PO", "TPP", "POX", "CAT"))
boxplot.before.outliers
```

Remove outliers

```{r remove catalase outliers}
box.cat <- boxplot(combinedData$catNP, plot = FALSE)
combinedData$catNP[which(combinedData$catNP %in% box.cat$out)] <- "NA"
combinedData$catNP <- as.numeric(as.character(combinedData$catNP))
```

```{r remove peroxidase outliers}
box.pox <- boxplot(combinedData$poxN, plot = FALSE)
combinedData$poxN[which(combinedData$poxN %in% box.pox$out)] <- "NA"
combinedData$poxN <- as.numeric(as.character(combinedData$poxN))
```

```{r remove phenoloxidase outliers}
box.po <- boxplot(combinedData$poN, plot = FALSE)
combinedData$poN[which(combinedData$poN %in% box.po$out)] <- "NA"
combinedData$poN <- as.numeric(as.character(combinedData$poN))
```

```{r remove prophenoloxidase outliers}
box.ppo <- boxplot(combinedData$ppoN, plot = FALSE)
combinedData$ppoN[which(combinedData$ppoN %in% box.ppo$out)] <- "NA"
combinedData$ppoN <- as.numeric(as.character(combinedData$ppoN))
```


```{r outliers removed boxplot}
boxplot.removed.outliers <- combinedData %>% select(poN, ppoN, poxN, catNP, species) %>%
  pivot_longer(cols = c(poN, ppoN, poxN, catNP),
               names_to = "measurement",
               values_to = "value") %>%
  ggplot(aes(x = species, y = value, fill = measurement)) +
  facet_wrap(~ measurement, scales = "free_y") +
  geom_boxplot() +
  geom_jitter(width=0.05) +
  theme_classic() +
  scale_fill_manual(values=c("red", "green", "orange", "purple"),
                      name = "Measurement",
                    labels = c("PO", "TPP", "POX", "CAT"))
boxplot.removed.outliers
```

Save/Load

```{r save/load}
save(combinedData, file="immune_growth_data.RData")
load("immune_growth_data.RData")
```

# Investigate patterns between initial size and immune activity

Phenoloxidase ~ Initial Weight

```{r MCMC PO weight:species interaction raw}
set.seed(1)
mcmc.PO.wt.species.raw <- MCMCglmm(poN~WT1*species, 
                           random=~geno, 
                           data=combinedData)
summary(mcmc.PO.wt.species.raw)
```

```{r Plot PO by Initial Weight raw}
plotPO_WT1_raw <- combinedData %>%
  ggplot(aes(x=WT1, y=poN, color=species))+
  geom_point()+
  geom_smooth(method=lm, se=T, fullrange=TRUE)+
  theme_classic()+ 
  ylab("PO Activity")+
  xlab("Initial Weight (g)")+
  annotate("text", x = 12, y = 0.09, label = "W pMCMC < 0.001")+
  annotate("text", x = 12, y = 0.085, label = "S pMCMC = 0.534")+
  annotate("text", x = 12, y = 0.08, label = "W:S pMCMC = 0.166")+
  scale_color_manual(values = c("coral2", "steelblue1"),
                       name = "Species",
                       labels = c("M.cav", "O.fav"))
plotPO_WT1_raw
```

```{r MCMC PO  weight:species interaction log}
set.seed(1)
mcmc.PO.wt.species <- MCMCglmm(log(poN,2)~log(WT1,2)*species, 
                           random=~geno, 
                           data=combinedData)
summary(mcmc.PO.wt.species)
```

```{r Plot PO by Initial Weight log}
plotPO_WT1 <- combinedData %>%
  ggplot(aes(x=log(WT1,2), y=log(poN,2), color=species))+
  geom_point()+
  geom_smooth(method=lm, se=T, fullrange=TRUE)+
  theme_classic()+ 
  ylab("Log2 PO Activity")+
  xlab("Log2 Initial Weight (g)")+
  annotate("text", x = 4.4, y = -9, label = "W pMCMC = 0.012")+
  annotate("text", x = 4.4, y = -9.5, label = "S pMCMC = 0.258")+
  annotate("text", x = 4.4, y = -10, label = "W:S pMCMC = 0.430")+
  scale_color_manual(values = c("coral2", "steelblue1"),
                       name = "Species",
                       labels = c("M.cav", "O.fav"))
plotPO_WT1
```


Total phenoloxidase potential ~ Initial Weight

```{r MCMC PPO weight:species interaction raw}
set.seed(1)
mcmc.PPO.wt.species.raw <- MCMCglmm(ppoN~WT1*species, 
                           random=~geno, 
                           data=combinedData)
summary(mcmc.PPO.wt.species.raw)
```

```{r Plot PPO by Initial Weight raw}
plotPPO_WT1_raw <- combinedData %>%
  ggplot(aes(x=WT1, y=ppoN, color=species))+
  geom_point()+
  geom_smooth(method=lm, se=T, fullrange=TRUE)+
  theme_classic()+ 
  ylab("PPO Activity")+
  xlab("Initial Weight (g)")+
  annotate("text", x = 12, y = 0.09, label = "W pMCMC < 0.001")+
  annotate("text", x = 12, y = 0.085, label = "S pMCMC = 0.234")+
  annotate("text", x = 12, y = 0.08, label = "W:S pMCMC = 0.094")+
  scale_color_manual(values = c("coral2", "steelblue1"),
                       name = "Species",
                       labels = c("M.cav", "O.fav"))
plotPPO_WT1_raw
```

```{r MCMC PPO weight:species interaction log}
set.seed(1)
mcmc.PPO.wt.species <- MCMCglmm(log(ppoN,2)~log(WT1,2)*species, 
                           random=~geno, 
                           data=combinedData)
summary(mcmc.PPO.wt.species)
```

```{r Plot PPO by Initial Weight log}
plotPPO_WT1 <- combinedData %>%
  ggplot(aes(x=log(WT1,2), y=log(ppoN,2), color=species))+
  geom_point()+
  geom_smooth(method=lm, se=T, fullrange=TRUE)+
  theme_classic()+ 
  ylab("Log2 PPO Activity")+
  xlab("Log2 Initial Weight (g)")+
  annotate("text", x = 4.4, y = -9, label = "W pMCMC = 0.006")+
  annotate("text", x = 4.4, y = -9.5, label = "S pMCMC = 0.888")+
  annotate("text", x = 4.4, y = -10, label = "W:S pMCMC = 0.942")+
  scale_color_manual(values = c("coral2", "steelblue1"),
                       name = "Species",
                       labels = c("M.cav", "O.fav"))
plotPPO_WT1
```

Peroxidase ~ Initial Weight

```{r MCMC POX weight:species interaction raw}
set.seed(1)
mcmc.POX.wt.species.raw <- MCMCglmm(poxN~WT1*species, 
                           random=~geno, 
                           data=combinedData)
summary(mcmc.POX.wt.species.raw)
```

```{r Plot POX by Initial Weight raw}
plotPOX_WT1_raw <- combinedData %>%
  ggplot(aes(x=WT1, y=poxN, color=species))+
  geom_point()+
  geom_smooth(method=lm, se=T, fullrange=TRUE)+
  theme_classic()+ 
  ylab("POX Activity")+
  xlab("Initial Weight (g)")+
  annotate("text", x = 12, y = 0.16, label = "W pMCMC < 0.001")+
  annotate("text", x = 12, y = 0.15, label = "S pMCMC = 0.754")+
  annotate("text", x = 12, y = 0.14, label = "W:S pMCMC = 0.532")+
  scale_color_manual(values = c("coral2", "steelblue1"),
                       name = "Species",
                       labels = c("M.cav", "O.fav"))
plotPOX_WT1_raw
```

```{r MCMC POX weight:species interaction log}
set.seed(1)
mcmc.POX.wt.species <- MCMCglmm(log(poxN,2)~log(WT1,2)*species, 
                           random=~geno, 
                           data=combinedData)
summary(mcmc.POX.wt.species)
```

```{r Plot POX by Initial Weight log}
plotPOX_WT1 <- combinedData %>%
  ggplot(aes(x=log(WT1,2), y=log(poxN,2), color=species))+
  geom_point()+
  geom_smooth(method=lm, se=T, fullrange=TRUE)+
  theme_classic()+ 
  ylab("Log2 POX Activity")+
  xlab("Log2 Initial Weight (g)")+
  annotate("text", x = 4.4, y = -11.8, label = "W pMCMC = 0.138")+
  annotate("text", x = 4.4, y = -12.2, label = "S pMCMC = 0.304")+
  annotate("text", x = 4.4, y = -12.6, label = "W:S pMCMC = 0.346")+
  scale_color_manual(values = c("coral2", "steelblue1"),
                       name = "Species",
                       labels = c("M.cav", "O.fav"))
plotPOX_WT1
```

Catalase ~ Initial Weight

```{r MCMC CAT weight:species interaction raw}
set.seed(1)
mcmc.CAT.wt.species.raw <- MCMCglmm(catNP~WT1*species, 
                           random=~geno, 
                           data=combinedData)
summary(mcmc.CAT.wt.species.raw)
```

```{r Plot CAT by Initial Weight raw}
plotCAT_WT1_raw <- combinedData %>%
  ggplot(aes(x=WT1, y=catNP, color=species))+
  geom_point()+
  geom_smooth(method=lm, se=T, fullrange=TRUE)+
  theme_classic()+ 
  ylab("CAT Activity")+
  xlab("Initial Weight (g)")+
  annotate("text", x = 12, y = 5.3, label = "W pMCMC < 0.001")+
  annotate("text", x = 12, y = 5, label = "S pMCMC = 0.040")+
  annotate("text", x = 12, y = 4.7, label = "W:S pMCMC = 0.110")+
  scale_color_manual(values = c("coral2", "steelblue1"),
                       name = "Species",
                       labels = c("M.cav", "O.fav"))
plotCAT_WT1_raw
```

```{r MCMC CAT weight:species interaction log}
set.seed(1)
mcmc.CAT.wt.species <- MCMCglmm(log(catNP,2)~log(WT1,2)*species, 
                           random=~geno, 
                           data=combinedData)
summary(mcmc.CAT.wt.species)
```

```{r Plot CAT by Initial Weight log}
plotCAT_WT1 <- combinedData %>%
  ggplot(aes(x=log(WT1,2), y=log(catNP,2), color=species))+
  geom_point()+
  geom_smooth(method=lm, se=T, fullrange=TRUE)+
  theme_classic()+ 
  ylab("Log2 CAT Activity")+
  xlab("Log2 Initial Weight (g)")+
  annotate("text", x = 4.4, y = -2.8, label = "W pMCMC < 0.001")+
  annotate("text", x = 4.4, y = -3, label = "S pMCMC = 0.080")+
  annotate("text", x = 4.4, y = -3.2, label = "W:S pMCMC = 0.086")+
  scale_color_manual(values = c("coral2", "steelblue1"),
                       name = "Species",
                       labels = c("M.cav", "O.fav"))
plotCAT_WT1
```

