---
title: "Microfragmentation Growth Data"
author: "Lou Schlecker & Rachel Wright"
date: "7/27/2021"
output:
  html_document: default
  pdf_document: default
---

```{r load libraries, include = FALSE}
library(tidyverse) # for data wrangling and plotting
library(MCMCglmm) # for Bayesian generalized linear mixed model statistical analysis

# setwd("Z:/Lou_Fragmentation/R_test/growth_data_ready/")
```

# Growth Data Analysis
We have growth data for 152 microfragments

```{r load growth data}
growthData <- read.csv("growthData.csv")
head(growthData)
```

Combine Genets X and V, which we know from sequence data are clones
Rename them both "U"

```{r Combine clones growth data}
growthData$geno <- gsub("X|V", "U", growthData$geno)
growthData$geno
```

Add species information
Genets at the beginning of the alphabet = Montastraea cavernosa
Genets at the end of the alphabet = Orbicella faveolata

```{r add species}
mcav <- c("A", "B", "C", "D", "E")
ofav <- c("U", "W", "Y", "Z")

growthData <- growthData %>% 
  mutate(species = ifelse(geno %in% mcav, "mcav", "ofav"))
```

Explore growth data

```{r histogram weight data raw}
growthData %>% select(species, WT1, WT2) %>%
  pivot_longer(cols = c(WT1, WT2),
               names_to = "measurement",
               values_to = "value") %>%
  ggplot(aes(x = value, fill = measurement)) +
  facet_grid(~ species) +
  ylab("Count") +
  xlab("Weight (g)") +
  geom_histogram(position = "identity", alpha = 0.7, bins = 20) +
  theme_classic() +
  scale_fill_manual(values=c("coral2", "dodgerblue1"),
                      name = "Measurement",
                    labels = c("Initial Weight", "Final Weight"))
```

```{r histogram weight data log2}
growthData %>% select(species, WT1, WT2) %>%
  pivot_longer(cols = c(WT1, WT2),
               names_to = "measurement",
               values_to = "value") %>%
  ggplot(aes(x = log(value,2), fill = measurement)) +
  facet_grid(~ species) +
  ylab("Count") +
  xlab("Log2 weight (g)") +
  geom_histogram(position = "identity", alpha = 0.7, bins = 20) +
  theme_classic() +
  scale_fill_manual(values=c("coral2", "dodgerblue1"),
                      name = "Measurement",
                    labels = c("Initial Weight", "Final Weight"))
```

```{r histogram area data raw}
growthData %>% select(species, area, areaFinal) %>%
  pivot_longer(cols = c(area, areaFinal),
               names_to = "measurement",
               values_to = "value") %>%
  ggplot(aes(x = value, fill = measurement)) +
  facet_grid(~species) +
  ylab("Count") +
  xlab("Area (cm2)") +
  geom_histogram(position = "identity", alpha = 0.7, bins = 20) +
  theme_classic() +
  scale_fill_manual(values=c("coral2", "dodgerblue1"),
                      name = "Measurement",
                    labels = c("Initial Area", "Final Area"))
```

```{r histogram area data log}
growthData %>% select(species, area, areaFinal) %>%
  pivot_longer(cols = c(area, areaFinal),
               names_to = "measurement",
               values_to = "value") %>%
  ggplot(aes(x = log(value,2), fill = measurement)) +
  facet_grid(~species) +
  ylab("Count") +
  xlab("Log2 area (cm2)") +
  geom_histogram(position = "identity", alpha = 0.7, bins = 20) +
  theme_classic() +
  scale_fill_manual(values=c("coral2", "dodgerblue1"),
                      name = "Measurement",
                    labels = c("Initial Area", "Final Area"))
```

Calculate log-fold change growth

```{r log-fold change in WT}
growthData <- growthData %>% mutate(logWTchange = log((WT2/WT1),2))
```

```{r hist log fold change WT}
growthData %>%
  ggplot(aes(x = logWTchange, fill = species)) +
  ylab("Count") +
  xlab("Log-fold Weight Change") +
  geom_histogram(position = "identity", alpha = 0.5, bins = 20) +
  theme_classic() +
  scale_fill_manual(values=c("coral2", "dodgerblue1"),
                      name = "Measurement",
                    labels = c("M. cavernosa", "O. faveolata"))
```


Microfragmentation growth analysis
Do smaller fragments grow faster?

Compare initial weight (g) to percent weight change

```{r MCMC growth initial weight:species interaction}
set.seed(1)
mcmc.percWT <- MCMCglmm(log(percGrowth,2)~log(WT1,2)*species, 
                           random=~geno, 
                           data=growthData)
summary(mcmc.percWT)
```

```{r Plot Percent Growth by Initial Weight}
percGrowthbyInitialWeight <- growthData %>%
  ggplot(aes(x=log(WT1,2), y=log(percGrowth,2), color=species))+
  geom_point()+
  geom_smooth(method=lm, se=T, fullrange=TRUE)+
  theme_classic()+ 
  ylab("Log 2Percent Weight Increase")+
  xlab("Log2 Initial Weight")+
  annotate("text", x = 4, y = 1.2, label = "W pMCMC < 0.001")+
  annotate("text", x = 4, y = 1.0, label = "S pMCMC = 0.396")+
  annotate("text", x = 4, y = 0.8, label = "W:S pMCMC = 0.712")+
  scale_color_manual(values = c("coral2", "dodgerblue1"),
                       name = "Species",
                       labels = c("M.cav", "O.fav"))
percGrowthbyInitialWeight
```




Compare area (cm2) to percent area change

```{r MCMC growth area:species interaction}
set.seed(1)
mcmc.percArea <- MCMCglmm(log(percArea,2)~log(area,2)*species, 
                           random=~geno, 
                           data=growthData)
summary(mcmc.percArea)
```

```{r Plot Percent Area by Initial Area}
percGrowthbyInitialArea <- growthData %>%
  ggplot(aes(x=log(area,2), y=log(percArea,2), color=species))+
  geom_point()+
  geom_smooth(method=lm, se=T, fullrange=TRUE)+
  theme_classic()+ 
  ylab("Log 2Percent Area Increase")+
  xlab("Log2 Initial Area")+
  annotate("text", x = -2, y = 1.4, label = "A pMCMC < 0.001")+
  annotate("text", x = -2, y = 1.1, label = "S pMCMC = 0.150")+
  annotate("text", x = -2, y = 0.8, label = "A:S pMCMC = 0.032")+
  scale_color_manual(values = c("coral2", "dodgerblue1"),
                       name = "Species",
                       labels = c("M.cav", "O.fav"))
percGrowthbyInitialArea
```

Compare initial weight to log-fold weight change

```{r MCMC growth initial weight:species interaction log fold}
set.seed(1)
mcmc.logWT <- MCMCglmm(logWTchange~log(WT1,2)*species, 
                           random=~geno, 
                           data=growthData)
summary(mcmc.logWT)
```

## New log-fold change by weight analysis

```{r Plot Log Growth by Initial Weight}
logGrowthbyInitialWeight <- growthData %>%
  ggplot(aes(x=log(WT1,2), y=logWTchange, color=species))+
  geom_point()+
  geom_smooth(method=lm, se=T, fullrange=TRUE)+
  theme_classic()+ 
  ylab("Log2 Weight Change (WTfinal/WTinitial)")+
  xlab("Log2 Initial Weight")+
  annotate("text", x = 6, y = 0.25, label = "W pMCMC < 0.001")+
  annotate("text", x = 6, y = 0.23, label = "S pMCMC = 0.194")+
  annotate("text", x = 6, y = 0.21, label = "W:S pMCMC = 0.404")+
  scale_color_manual(values = c("coral2", "dodgerblue1"),
                       name = "Species",
                       labels = c("M.cav", "O.fav"))
logGrowthbyInitialWeight
```

What size is best? Assume you have 100 g of starting material and you fragment them into smaller pieces.

```{r}
max(growthData$WT1)

growthData <- growthData %>%
  mutate(growth_norm_max_size = (WT2-WT1)*(100/WT1))

growthData %>% ggplot(aes(x = WT1, y = growth_norm_max_size, color = species)) +
  geom_point() +
  stat_smooth(method = "loess", formula = y ~ x, size = 1) +
  facet_wrap(~ species, scales="free") +
  scale_color_manual(values = c("coral2", "dodgerblue1"),
                       name = "Species",
                       labels = c("M.cav", "O.fav")) +
  ylab("Total Predicted Growth (g) in 4 Months") +
  xlab("Starting Size (g)") +
  theme_classic()
```

What size is best? Assume you have 100 cm2 of starting material and you fragment them into smaller pieces.


```{r}
max(growthData$area)

growthData <- growthData %>%
  mutate(area_norm_max_size = (areaFinal-area)*(100/area))

growthData %>% ggplot(aes(x = area, y = area_norm_max_size, color = species)) +
  geom_point() +
  stat_smooth(method = "loess", formula = y ~ x, size = 1) +
  facet_wrap(~ species, scales="free") +
  scale_color_manual(values = c("coral2", "dodgerblue1"),
                       name = "Species",
                       labels = c("M.cav", "O.fav")) +
  ylab("Total Predicted Growth (cm2) in 4 Months") +
  xlab("Starting Size (cm2)") +
  theme_classic()
```

What size is best? Plot growth without log transformation

```{r}
growthData %>% ggplot(aes(x = area, y = percArea, color = species)) +
  geom_point() +
  stat_smooth(method = "loess", formula = y ~ x, size = 1) +
  facet_wrap(~ species, scales="free") +
  scale_color_manual(values = c("coral2", "dodgerblue1"),
                       name = "Species",
                       labels = c("M.cav", "O.fav")) +
  ylab("Percent Growth Rate (Final-Initial/Initial*100)") +
  xlab("Starting Size (cm2)") +
  geom_vline(xintercept=1, linetype="dotted")+
  theme_classic()

min(growthData$percArea)
max(growthData$percArea)
```

## Association between growth and proportion clade D in M. cav

```{r}
symFracs <- read.delim("allcounts_symABCD.txt")
symFracs$sample <- substr(symFracs$sample,1,3)
symFracs$sample <- gsub("-","",symFracs$sample)
symFracs$sample <- gsub("_","",symFracs$sample)
```

Average technical replicates

```{r}
symFracs0 <- symFracs %>% group_by(sample) %>% mutate(meanFracD = mean(fracD)) %>%
  select(sample, meanFracD) %>% filter(!duplicated(sample))
```

```{r}
growthD <- merge(growthData, symFracs0, by.x="sam", by.y="sample")
```

```{r MCMC growth area:fracD}
set.seed(1)
mcmc.areaD <- MCMCglmm(percArea~meanFracD, 
                           random=~geno, 
                           data=growthD)
summary(mcmc.areaD)
```

```{r Plot Area by Frac D 1}
percAreabyFracD <- growthD %>%
  ggplot(aes(x=meanFracD, y=percArea))+
  geom_point()+
  geom_smooth(method=lm, se=T, fullrange=TRUE)+
  theme_classic()+ 
  ylab("Percent Area Increase")+
  xlab("Fraction Clade D")+
  annotate("text", x = 0.75, y = 200, label = "pMCMC = 0.068")
percAreabyFracD
```

```{r MCMC growth weight:fracD}
set.seed(1)
mcmc.weightD <- MCMCglmm(percGrowth~meanFracD, 
                           random=~geno, 
                           data=growthD)
summary(mcmc.weightD)
```

```{r Plot Area by Frac D}
percWeightbyFracD <- growthD %>%
  ggplot(aes(x=meanFracD, y=percGrowth))+
  geom_point()+
  geom_smooth(method=lm, se=T, fullrange=TRUE)+
  theme_classic()+ 
  ylab("Percent Buoyant Weight Increase")+
  xlab("Fraction Clade D")+
  annotate("text", x = 0.75, y = 10, label = "pMCMC = 0.422")
percWeightbyFracD
```


